# STATS 765, Lab 01

Use {tidyverse} for data wrangling and insight exploration tasks.

### Task 1: Generate alarms data

The following code snippet was used to generate the mock-up alarm frequency and reginal information, which that were used as examples in the lecture last week. Run the code with your student ID as seed variable and generate your own (random) data

```{r task1, include=TRUE}
# load library
library(tidyverse)

# set random seed
seed <- 991 # <Replace '765' with your student ID here.>
set.seed(seed)

# generate alarm id, dates and region vectors
# paste0: concatenate string (0 means there is no space)
alarms_id.vt <- paste0('alarms_', sample(LETTERS, size = 10, replace = FALSE))
alarms_dates.vt <- paste0('d.',seq(Sys.Date()-60, Sys.Date(), by = '1 day'))  # 60 days back.
alarms_region.vt <- c('AKL_North', 'AKL_Central', 'Waiheke','AKL_South', 'AKL_Others')

# generate random alarm frequency counts
alarms_count.mt <- matrix(round(runif(length(alarms_id.vt) * length(alarms_dates.vt))*seed),
                          nrow = length(alarms_id.vt), ncol=length(alarms_dates.vt))
colnames(alarms_count.mt) <- alarms_dates.vt

# set up data frames
alarms_count.df <- data.frame(alarm_id = alarms_id.vt, alarms_count.mt)
alarms_info.df <- data.frame(alarm_ID = alarms_id.vt, alarms_region = alarms_region.vt)
```

### Task 2: {tidyverse} Operations

Use the data generated to answer the following question: what is the average number of alarms per day in each region?

```{r task2, include=TRUE}
alarms_insight_df <- alarms_count.df %>%
  # pivot date data for joining with info data frame
  gather('date', 'frequency', -alarm_id) %>%
  # separate out dates
  separate(date, c('prefix', 'year', 'month', 'day')) %>%
  # remove prefix column
  select(-prefix) %>%
  # rejoin dates
  left_join(alarms_info.df, by = c('alarm_id' = 'alarm_ID')) %>%
  # clean up col order
  group_by(year, month, day, alarms_region) %>%
  summarise(average_count = round(mean(frequency),2))

knitr::kable(head(alarms_insight_df, n = 10))
```

# Task 3: Date operation using {lubridate}

Use the data generated to answer the following question: which days of the week has the 
highest and the lowest average number of alarms across the entire Auckland?

```{r task3, include=TRUE}
library(lubridate)

alarms_insight_df <- alarms_count.df %>%
  gather('date', 'frequency', -alarm_id) %>%
  # date is currently a string, need to convert to datetime object
  mutate(dd = as.Date(substring(date, 3), format = '%Y.%m.%d')) %>%
  mutate(week_day = wday(dd, label = TRUE, abbr = TRUE)) %>%
  # clean up
  select(-dd, -date) %>%
  left_join(alarms_info.df, by = c('alarm_id' = 'alarm_ID')) %>%
  group_by(week_day) %>%
  summarise(average_count = round(mean(frequency),2)) %>%
  arrange(average_count)
```

### Task 4: Retrospective

Write a short paragraph summarising your experience 
and comments about using {tidyverse} for data wrangling tasks.  

I really enjoyed learning the most crucial data wrangling functions in R {tidyverse} for data science
Wrangling is a process which I would like to get more understanding on. In particular what is best
practice on order of operations i.e what do we clean up first.
I believe it's the most crucial step in data science as no science can be done with 'dirty' data
Hopefully data wrangling will click to me over the next few weeks!
