# STATS 765, Lab 03

## Lab 03 Overview

In the last lecture, we explored the sandwich estimator, which allows us to make valid estimation of se(β̂ )
 for linear regression, without assuming the usual constant-variance of the residual errors.

This lab is to gain a better understanding of this idea with a simulated data set.

## Task 1: Generate data
To illustrate the idea of using sandwich estimator, we first generate a set of data with increased variances as the predictor.

### Step 1.1: Run the following piece of code, with your student as seed.

```{r}
set.seed(991) # replace "765" with your student ID.
n = 200  
x = rnorm(n)
residual_std = exp(x) # error standard deviation is exponential w.r.t. x values, 
y = 1.5+3*x + residual_std*rnorm(n)
```

### Step 1.2: Plot y vs. x and comment

```{r}
# create dataframe
x_y_df = data.frame(cbind(x,y))
plot(y ~ x, xlab = "x", ylab = "y", data = x_y_df)
```


There is a clear linear trend between $x$ and $y$.
There is non-constant scatter, as the scatter increases as the value of x increases

## Task 2: Fit the linear regression model as usual

### Step 2.1: Fit a linear regression model

Fit a linear regression model y~x and generate the model summary output.

```{r}
x_y_fit = lm(y ~ x, data = x_y_df)
summary(x_y_fit)

plot(y ~ x, xlab = "x", ylab = "y", data = x_y_df)
abline(coef(x_y_fit)[1], coef(x_y_fit)[2])
```

### Step 2.2: Infererence based on summary output

What is the standard error of coefficient of X? What is the 95% confidence interval of the β1^
? Interprete the p-value shown.

- Standard error for $\hat{\beta_1}$ is `r round(summary(x_y_fit)$coef["x", "Std. Error"], 3)`.\
- The p-value tests the beta values to the null hypothesis. The null hypothesis being that the beta values are equal to 0. We have very strong evidence (P value = `r options(digits=3); summary(x_y_fit)$coef["x", "Pr(>|t|)"]`, we can reject the null hypothesis and make an inference that x and y are independent.\
- The 95% confidence interval is `r confint(x_y_fit)["x", "2.5 %"]` and `r confint(x_y_fit)["x", "97.5 %"]`.\
- t-statistic is calculated by the estimate minus the null hypothesis, divided by the standard error. We use the t-statistic to calculate the p-value.\

## Task 3: Sandwich estimator

### Step 3.1: Calculate se(βi^) use the sandwich estimator.

```{r}

```

 
### Step 3.2: Inference

Use the sandwich standard error for the cofficient of X to calculate the 95% confidence interval and p-value for β1

# Task 4: Comparison and Comment

What can you observe from comparing the two sets of inference results above?



